---
title: "Kobe Bryant Shooting"
author: "Ryan Burge"
date: "April 27, 2016"
output: html_document
---

# Kobe Bryant's Success Against Different Opponents and Throughout His Career


```{r message=FALSE, warning=FALSE}
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function
library(dplyr)
library(DT)
```

```{r message=FALSE, warning=FALSE}
kobe <- read.csv("D:/kobe/kobe.csv", stringsAsFactors = FALSE)
```

I want to do some quick and broad data exploration to see if Kobe is a better shooter against certain opponents. 

I need to take a look at the important variable first, kobe$shot_made_flag.

```{r message=FALSE, warning=FALSE}
head(kobe$shot_made_flag, 50)
```

There are a lot of NAs in there, that will make this difficult so I am just going to ignore them for this analysis, I don't think it can make a huge difference in the final analysis. 

```{r message=FALSE, warning=FALSE}
team <- aggregate(kobe$shot_made_flag, list(kobe$opponent), na.rm = TRUE, mean)
team <- arrange(team, x)
head(team, 33)
```
So, I've got what I need now. A smaller dataframe with just the opponents name and Kobe's shooting percentage against them. I want to add a little extra analysis, so I am going to create a vector that lists the conferences of each opponent. There's not an easy way to do this, so I will just do it by hand. 

```{r message=FALSE, warning=FALSE}
team$conference <- c("Eastern", "Eastern", "Western", "Eastern", "Eastern" ,"Western", "Eastern", "Eastern" , "Eastern" , "Western" , "Eastern", "Eastern", "Eastern", "Eastern" , "Western" , "Eastern", "Eastern", "Western", "Western", "Eastern", "Western" , "Eastern", "Eastern", "Western", "Western", "Western", "Western" , "Western", "Western", "Western", "Western", "Western", "Eastern" )
```
Looks good so let's plot. I want this to be plotted in order of best shooting percentage to worst shooting percentage. I am also going to add in a horizontal line indicating Kobe's overall shooting percentage, which is 43.9%. 

```{r message=FALSE, warning=FALSE}
ggplot(team, aes(x=reorder(Group.1, -x), y = x)) + geom_bar(aes(fill=conference),stat="identity")  + ylim(0,.5) + theme(axis.text.x = element_text(angle = 90)) + xlab("Opponent's Name") + ylab("Shooting Percentage") + geom_hline(yintercept = .43)
```

So there's something interesting. Kobe shoots better against the Western conference than the Eastern Conference. That may be an interesting feature to use. 

I also want to see if Kobe really declined near the end of his career. I know that he tore his Achilles and missed a lot of time. Again, I'm adding a line to indicate Kobe's lifetime shooting percentage. 

```{r message=FALSE, warning=FALSE}
season <- aggregate(kobe$shot_made_flag, list(kobe$season), na.rm = TRUE, mean)
ggplot(season, aes(x=Group.1, y = x)) + geom_bar(stat="identity")  + ylim(0,.5) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + xlab("Season") + ylab("Shooting Percentage") + geom_hline(yintercept = .43)
```

Yeah, it's pretty clear that Kobe was really consistent for a long period of time. He was a very consistent shooter from 1996 to 2013-2014. The last three years were ugly with his shooting percentage doropping pretty significantly near the end. 

Here's a quick data table to show that decline in a different format.  
```{r message=FALSE, warning=FALSE}
datatable(season, class = 'compact')
```


Those last three years are pretty brutal. 40.6%, 37.6%, and 35.6%. Much lower than Kobe was used to. He knew he was in terminal decline. 

